---
title: "HF2_final"
author: "B치n칩czi Anna, Heilmann Istv치n, Reizinger Krist칩f"
date: '2019 11 19 '
output: html_document
---

#Required packages
```{r}
library(assertthat)
library(corrplot)
library(zoo)
```


#The first function checks whether readxl package is installed on your computer or not
#This is a side function for import_dataset_WTI


```{r}
pack_check <- function()
{
  my_packages <- library()$results
  n = length(my_packages)
  readpackage = 0
  for(i in 1:n)
  {
    if (my_packages[i] == "readxl")
    {
      readpackage <- i
    }
  }
  return(readpackage)
}
```



##The next function imports the (WTI) dataset that we gave in and then shows the database

```{r}
import_dataset_WTI <- function()
{
   path_wti <- "C:/Users/banna/Documents/egyetem/EFRP/R//WTI2.xlsx"
  if (pack_check() == 0)
  {
    # readxl package is not installed, so now we have to
    install.packages("readxl")
  }
  library("readxl")
  WTI_fut <<- read_excel(path_wti)
  # View(WTI_fut)
  return(WTI_fut)
}

```

#The next function converts the raw dataset into a new one by making 
#The returns from day n to day n + 1
#The first column remains the same as it is our date vector, 
#And the first row is empty as we don't have data before the first day

```{r}
return_maker <- function(WTI_fut)
{
  n <- nrow(WTI_fut)
  m <- ncol(WTI_fut)
  ret_WTI_fut <<- matrix(nrow = n, ncol = m)
  for (i in 2:n)
  {
    for (j in 2:m)
    {
      ret_WTI_fut[i,j] <<- (WTI_fut[[i,j]] / WTI_fut[[i-1,j]]) - 1
    }
  }
  
  ret_WTI_fut <<- cbind(WTI_fut[,1], ret_WTI_fut)
  ret_WTI_fut = ret_WTI_fut[-1,-2]
  colnames(ret_WTI_fut) = colnames(WTI_fut)
  # View(ret_WTI_fut)
  return(ret_WTI_fut)
}

```


#This function's inputs will be saved for other functions's parameters
#Please give the dates in this format : "yyyy-mm-dd"

```{r}
parameter_lists <- function(asset1, asset2, startDate, endDate, windowLength, lag)
{
  parameters<-list(asset1,asset2,startDate,endDate,windowLength,lag)
  return(parameters)
  
}

```

#This function generates the correlation-timeseries table
#The column name's first part shows the unlagged vector,the second part shows the lagged one
#Please type the dates in this format : "yyyy-mm-dd"
#Lag can be negative or positive, depends on the direction you want. 
#e.g. -5 lagg means that the other vector will be lagged back in time
#Asset1 and asset2 are integers

```{r}
time_series_generating <- function(parameters)
{
  asset1<-parameters[[1]]
  asset2<-parameters[[2]]
  startDate<-parameters[[3]]
  endDate<-parameters[[4]]
  windowLength<-parameters[[5]]
  lag<-parameters[[6]]
  
  
  absStart <- min(ret_WTI_fut[,1]) #first date in the vector
  absEnd <- max(ret_WTI_fut[,1]) #largest date in the vector
  
  if (as.Date(startDate, origin = "1970-01-01") > as.Date(endDate, origin = "1970-01-01"))
  {
    return("StartDate starts after EndDate!")
  }
  if (as.Date(startDate, origin = "1970-01-01") + lag < as.Date(absStart, origin = "1970-01-01"))
  { 
    return("The starting date and the lag gets out of our dataset!")
  }
  if (as.Date(endDate, origin = "1970-01-01") + lag > as.Date(absEnd, origin = "1970-01-01")) 
  {
    return("The ending date and the lagg gets out of the dataset!")
  }
  
  for (i in 1:length(asset1))
  {
    if (as.Date(ret_WTI_fut[[i,1]]) - as.Date(startDate, origin = "1970-01-01") == 0)
    {
      startDateIndex = i 
    }
  }
  
#first row if return matrix is NA for each item!
  n <- (as.numeric(as.Date(endDate), origin = "1970-01-01")) - (as.numeric(as.Date(startDate), origin = "1970-01-01")) - windowLength 
#n is the length of the time series
  
  
  correlations <- vector(length = n)
  timeVector <- vector(length = n)
#calculate the values of the vector for each item
  for (i in 1:n)
  {
    correlations[i] = cor(asset1[(startDateIndex + i - 1):(startDateIndex + i + windowLength - 1)],
                          asset2[(startDateIndex - 1 + i + lag):(startDateIndex - 1 + i + lag + windowLength)])
    timeVector[i] = as.Date(as.numeric(as.Date(startDate), origin="1970-01-01") + i - 1, origin = "1970-01-01")
  }
  
  class(timeVector) <- "Date"
  
  timeSeries <- data.frame(timeVector, correlations)
  return(timeSeries)
}

```



#New object in order to have a network

```{r}
network_statistics <-function(startdate=sdate, enddate=edate, ws=winsize,cl=10)
{
# parameters
# startdate first date of the analyzed period
# enddate last date of the analyzed period
# windowsize
# cl correlation lag
  
second_dim <- as.numeric(as.Date(enddate) - as.Date(startdate)) - ws
# length of the

m <-24 # number of assets

# select dates from the original data table
from <- which(as.Date(ret_WTI_fut[,1])==startdate) # first coordinate
to <- which(as.Date(ret_WTI_fut[,1])==enddate)-ws-1 # second coordinate
dates <-zoo::as.Date(ret_WTI_fut[-1,1][from:to]) # dates vector

dates <- as.character(dates) # labels for columns

assets <- as.character(seq(1,m)) # assets indicators from 1 to 24 (m=24)

network <- array(dim = c(m, second_dim, m), data = NA, dimnames = list(assets,dates, assets))
# 3 dimensional array for calculations

# calxulation of cross-correlations for all date sand assets
for (i in 3:(m+2)) {
  for (j in 3:(m+2)) {
    # tempt <- time_series_generating(parameter_lists(ret_WTI_fut[-1,i], ret_WTI_fut[-1,j], startdate, enddate, ws,cl))[,2]
    tempt <- time_series_generating(parameter_lists(ret_WTI_fut[-1,i], ret_WTI_fut[-1,j], startdate, enddate, ws,cl))[,2]
    
    for (k in 1:second_dim) {
      network[i-2, k, j-2] <- tempt[k]

    }
  }
}

return(network)

}

```


#Making the plots of the time series
```{r}
plot_timeseries <- function(network=network,startdate=sdate,enddate=edate, ws=winsize )
{
  second_dim <- as.numeric(as.Date(enddate) - as.Date(startdate)) - ws
  
  networkmmm <- matrix(nrow = second_dim,ncol=3, data = NA) # matrix for Min, Max, Mean
  colnames(networkmmm) <- c("min", "max", "mean")
  for(l in 1:second_dim){
    networkmmm[l,"min"] <- min(network[,l,])
    networkmmm[l,"max"] <- max(network[,l,])
    networkmmm[l,"mean"] <- mean(diag(network[,l,]) + network[,l,])/2
  }

  ts1 <- networkmmm[,"min"]
  ts2 <- networkmmm[,"max"]
  ts3 <- networkmmm[,"mean"]
  
  from <- which(as.Date(ret_WTI_fut[,1])==startdate)
  to <- which(as.Date(ret_WTI_fut[,1])==enddate)-ws-1
  dates <-zoo::as.Date(ret_WTI_fut[-1,1][from:to])

  plot(dates, ts1, "l", col = "red", xlab = "Time", ylab = "Correlation", main = "Dynamic cross-correlations")

  lines(dates, ts2, "l", col = "blue")

  lines(dates, ts3, "l", col="green")

  legend("bottomleft", legend = c("min","max","mean"), fill = c("red","blue","green"), cex = 0.6)
  
  
}
```

#For plotting the correlations
```{r}
corr_plot <- function(network=network, day="2013-06-10" ){
  assertthat::assert_that(zoo::as.Date(day)>zoo::as.Date(sdate),msg=paste0("Choose a date another date!","Day must be later than ", sdate, " !"))
  assertthat::assert_that(zoo::as.Date(day)<zoo::as.Date(edate),msg=paste0("Choose a date another date!","Day must be earlier than ", sedate, " !"))
  
corrplot::corrplot(network[,day,], method = "square", cl.lim = c(-0.2, 0.2), 
         col = colorRampPalette(c("blue", "green"))(100))
 
}
```

#Network plot between assets pairwise cross-correlations
#Trying out the functions

```{r}
pack_check()

import_dataset_WTI()

return_maker(WTI_fut)
```


# Inputs for network and plotting

```{r}

sdate <- "2012-01-01"
edate <- "2014-12-31"
winsize <- 100
```

#Plots
```{r}
nw <-network_statistics(startdate=sdate, enddate=edate, ws=winsize, cl=10)
plot_timeseries(network=nw,startdate=sdate,enddate=edate, ws=winsize)
corr_plot(network=nw, day="2013-06-10")
```

